---
title: "Final Proposal"
author: "Emma Jay"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup}
library(tidyverse)

# set default theme for ggplot2
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 16))

# set default figure parameters for knitr
knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = 0.618,
  fig.retina = 2,
  dpi = 150, 
  out.width = "70%"
)
```

## The Dataset: The Great American Coffee Taste Test
```{r data load and cleaning}
# load data
raw_coffee_data <- read_csv("data/coffee_survey.csv")
glimpse(raw_coffee_data)

# clean and organize columns
coffee_data_clean <- raw_coffee_data %>%
  mutate(across(where(is.character), ~na_if(., "NA"))) %>%
  mutate(across(where(is.character), ~na_if(., "")))

coffee_data_clean <- coffee_data_clean %>%
  mutate(
    total_spend = as.numeric(total_spend)) %>%
  mutate(across(c
                (political_affiliation, 
                  education_level, 
                  ethnicity_race, 
                  gender, 
                  employment_status,
                  age,
                  cups),
                ~factor(.))) %>%
  mutate(
    brew = str_to_lower(brew),
    where_drink = str_to_lower(where_drink))

# remove predominately NA columns
prop_missing <- sapply(coffee_data_clean, function(x) mean(is.na(x)))
print(prop_missing)
coffee_data_clean <- coffee_data_clean %>%
  select(where(~mean(is.na(.)) < 0.9))

# check dataset
str(coffee_data_clean)
head(coffee_data_clean)
```
### Description of Data

## Research Question
Are there significant differences in coffee consumption patterns between individuals of different political affiliations within the same education or ethnic group?

## Graph 1
```{r graph 1}

```

### Discussion of Graph 1 (200 words)

## Draft Graph 2

## Draft Graph 3

## Draft Graph 4

### Discussion of Graphs 2 - 4 (400ish)
Along with these graphs, I would like to create a Shiny Dashboard that would allow a viewer to select specific education levels or ethnic groups that would update visualizations to be able to further explore the visualizations above. I explored this type of dashboard in Tableau, and I really enjoyed the interactive element it incorporated into the dashboard so I would like to incorporate something similar into my final project through Shiny. 
