---
title: "Final Proposal"
author: "Emma Jay"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup}
library(tidyverse)
library(plotly)
library(ggmosaic)

# set default theme for ggplot2
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 16))

# set default figure parameters for knitr
knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = 0.618,
  fig.retina = 2,
  dpi = 150, 
  out.width = "70%"
)
```

## The Dataset: The Great American Coffee Taste Test
```{r data load and cleaning}
# load data
raw_coffee_data <- read_csv("data/coffee_survey.csv")
glimpse(raw_coffee_data)

# clean and organize columns
coffee_data_clean <- raw_coffee_data %>%
  mutate(across(where(is.character), ~na_if(., "NA"))) %>%
  mutate(across(where(is.character), ~na_if(., "")))

coffee_data_clean <- coffee_data_clean %>%
  mutate(
    total_spend = as.numeric(total_spend)) %>%
  mutate(across(c
                (political_affiliation, 
                  education_level, 
                  ethnicity_race, 
                  gender, 
                  employment_status,
                  age),
                ~factor(.))) %>%
  mutate(
    brew = str_to_lower(brew),
    where_drink = str_to_lower(where_drink)) %>%
  mutate(cup_num = case_when(
    cups == "Less than 1" ~ 0,
    cups == "1" ~ 1,
    cups == "2" ~ 2,
    cups == "3" ~ 3,
    cups == "More than 4" ~ 5,
    cups == "4" ~ 4,
    TRUE ~ NA_real_
  ))

distinct(raw_coffee_data, cups)
# remove predominately NA columns
prop_missing <- sapply(coffee_data_clean, function(x) mean(is.na(x)))
print(prop_missing)
coffee_data_clean <- coffee_data_clean %>%
  select(where(~mean(is.na(.)) < 0.9))

# check dataset
str(coffee_data_clean)
head(coffee_data_clean)


```
### Description of Data

## Research Question
Are there significant differences in economic coffee consumption patterns between individuals of different political affiliations within the same education or ethnic group?

## Graph 1
```{r graph 1 - cups political and education}
cups_poli_edu <- ggplot(coffee_data_clean, aes(
  x = political_affiliation, 
  y = cup_num, 
  fill = political_affiliation)) +
  geom_boxplot() +
  facet_wrap(~ education_level) +
  labs(title = "Coffee Cups by Political Affiliation and Education Level",
       x = "Political Affiliation",
       y = "Cups of Coffee per Day",
       fill = "Political Affiliation") +
  theme_minimal()

ggplotly(cups_poli_edu)
```

### Discussion of Graph 1 (200 words)

## Draft Graph 2
```{r cafe value politic ethnicity}
convert_most_paid <- function(value) {
  case_when(
    value == "Less than $2" ~ 1,
    value == "More than $20" ~ 22,
    TRUE ~ {
      nums <- as.numeric(unlist(
        str_extract_all(value, "[0-9]+")))
      if (length(nums) == 2) {
        mean(nums) 
      } else {
        NA_real_
      }
    }
  )
}
coffee_data_clean <- coffee_data_clean %>%
  mutate(most_paid_numeric = sapply(
    most_paid, convert_most_paid))

ggplot(coffee_data_clean, aes(
  x = political_affiliation, 
  y = most_paid_numeric, 
  color = education_level)) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~ ethnicity_race)
  labs(title = "Cafe Value for Money by Political Affiliation and Ethnicity",
       x = "Political Affiliation",
       y = "Most Paid for Cup of Coffee",
       fill = "Education Level")
  
  distinct(coffee_data_clean, most_paid)
```

## Draft Graph 3
```{r }
distinct(coffee_data_clean, most_willing)

ggplot(coffee_data_clean, aes(
  x = most_willing, 
  fill = education_level)) +
  geom_bar(position = "dodge") +
  facet_wrap(~ political_affiliation) +
  labs(
    title = "Most Willing to Pay for a Cup of Coffee by Political Affiliation",
    x = "Most Willing Price Range",
    y = "Count",
    fill = "Education Level"
  ) 
```

### Discussion of Graphs 2 - 3
Along with these graphs, I would like to create a Shiny Dashboard that would allow a viewer to select specific education levels or ethnic groups that would update visualizations to be able to further explore the visualizations above. I explored this type of dashboard in Tableau, and I really enjoyed the interactive element it incorporated into the dashboard so I would like to incorporate something similar into my final project through Shiny. 

I think in regards to the draft graphs 2 and 3, I need to figure out better spacing or text size for the x-axis as the current format is very difficult to read. I like the idea of faceting these graphs across political affiliation or ethnic group, but I think when I consider their placement into a dashboard, I will need to think more critically about having multiple faceted visualizations and potentially look to visualize in such a way that does not need to be faceted or accounts for the interaction of the user as a means of separating out the interactions expressed by faceting above. Further graphs will definetly explore other economic related variables, as I am really interested in pulling out trends related to political affiliation, ethnicity, education and the idea of value or willingness to pay for coffee products.  
